---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{compstatr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Overview

## Getting Started with `R`

## Packages
In addition to `driftR`, you will also want to install and load other packages:

* [`dplyr`](http://dplyr.tidyverse.org) - a set of common verbs for cleaning and wrangling data
* [`readr`](http://readr.tidyverse.org) - tools for reading and writing plain-text data files
* [`tidyr`](http://tidyr.tidyverse.org) - 
* [`purrr`](http://purrr.tidyverse.org) - tools for working with functions and vectors
* [`lubridate`](http://lubridate.tidyverse.org) - tools for managing dates
* [`rlang`](https://rlang.r-lib.org/) - tools for working with core language features of R and tidyverse

these packages can be installed with the `install.packages()` function

## Using `Compastatr`
`cs_collapse()` - collapses multiple tibbles into one
`cs_combine()`- combines multiple data sets into one
`cs_crime()` - changes names of crimes to U.S. Federal Bureau of Investigation's Uniform Crime Reporting codes
`cs_crime_cat()` - Converts crime codes into strings
`cs_extract_month()`- Returns crime data of a selected month
`cs_filter_count()` - Removes counts of -1 from crime data
`cs_filter_crime()` - Returns crime data of a selected type of crime
`cs_is_int()` - Returns a logical vector stating if the reported address is an intersection
`cs_load_year()` - Loads the csv containing a year of crime data
`cs_missing_xy()` - Returns a logical vector stating if the reported address is missing coordinates
`cs_parse_date()` - Separates reported date and time into two columns 
`cs_parse_int()` - Separates intersections into two columns 
`cs_parse_month()`- Separates reported month and year into two columns 
`cs_replace_na()`- Replaces 0's with NA
`cs_standardize()` - Reformats nonstandard SLMPD time points into a 20 variable standard
`cs_validate_month()` - Identifies nonstandard variables within a month
`cs_validate_year()` - Ensures items are imported in the correct order, contain one month worth of data, and counts the numbers of variables. 
`january2018` - Contains SLMPD January 2018 crime data


### Importing Data
To import a year of crime data, `cs_load_year()`function is used. 

To use the package without importing your own data, you can load sample data in the package using:

```{r, eval=FALSE}
january2018 <- data("january2018")
```
```

### Function examples
This returns a logical vector containing TRUE in rows with violent crime
```{r, eval=FALSE}
violentCrimeSorted <- cs_crime(.data = january2018, var = Crime, newVar = ViolentCrime, crime = "violent"")
```
This returns a column containing the names, according to FBI crime codes, of crimes as a string
```{r, eval=FALSE}
crimeNames <- cs_crime_cat(.data = january2018, var = Crime, newVar = NamedCrime, output = string)
```
The data shows reported crimes and when multiple reports occur for the same indicent, the count is -1. This removes counts of -1. 
```{r, eval=FALSE}
january2018filteredCount <- cs_filter_count(.data = january2018, var = Count)
```
This returns a tibble containing only violent crime from the data set. 
```{r, eval=FALSE}
january2018violentCrimeFiltered <- cs_filter_crime(.data = january2018, var = Crime, crime = "violent")
```
Intersections are often reported in the same ILEADSStreet column and this creates a new column identifying if a crime was reported at an intersection. 
```{r, eval=FALSE}
january2018Intersections <- cs_filter_crime(.data = january2018, var = ILEADSStreet, newVar = "IsIntersection")
```

Some reported crimes present missing coordinates and this returns a logical vector with TRUE if a coordinate has no X and Y coordinates. 
```{r, eval=FALSE}
january2018MissingXY <- cs_missing_xy(.data = january2018, varx = XCoord, vary = YCoord, newVar = "IsMissingXandY")
```
This takes the reported and time and splits it into unique columns named "Date" and "Time" while keeping the input Date and time
```{r, eval=FALSE}
january2018ParsedDate <- cs_parse_date(.data = january2018, var = DateOccur, dateVar = Date, timeVar = Time, tz = Null, keepDateTime = TRUE)
```
This takes a column containing the month and a year and split it into two columns Year and Month
```{r, eval=FALSE}
january2018ParsedMonth <- cs_parse_month(.data = january2018, var = CodedMonth, newYear = Year, newMonth = Month)
```
In this example, missing X coordinates were inputted as 0 and the function replaces the missing coordinates with NA
```{r, eval=FALSE}
january2018XNA <- cs_replace_na(.data = january2018, var = XCoord)
```

### Cleaning one set of data example
